---
title: "Correlated tests statistics"
author: "Andrey Ziyatdinov"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    fig_caption: yes
    toc: true
    number_sections: true
vignette: >
  %\VignetteIndexEntry{Computation of linear models in matlm}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r options, echo = F}
opts_chunk$set(comment = NA, results = 'markup', tidy = F, message = F, warning = F, 
  echo = T, cache = T,
  fig.width = 6, fig.height = 6)
```

# About

## Links

- R script [demo/correlared-predictors.R](https://github.com/variani/matlm/blob/master/demo/correlared-predictors.R)


# Packages

```{r inc, cache = F}
library(devtools)
load_all("~/git/variani/matlm/")

load_all("~/git/variani/qq/")
```

```{r inc2, cache = F}
library(ggplot2)

theme_set(theme_light())
```

# Simulations parameters

```{r simpar}
N <- 2e3
M <- 2e3

seed <- 1

rho <- 0.9
```

# Independent predictors

```{r simpred_uncor, dependson = "simpar"}
simpred_uncor <- matlm_sim_randpred(seed = seed, N = N, M = M)
```

```{r assoc_uncor, dependson = -1}
assoc_uncor <- with(simpred_uncor, matlm(form, dat, pred = pred))
```

```{r qq_uncor}
qq_plot(assoc_uncor$tab$pval)
```

# Correlated predictors

```{r simpred_cor, dependson = "simpar"}
simpred_cor <- matlm_sim_randpred(seed = seed, N = N, M = M, rho = rho)
```

```{r assoc_cor, dependson = -1}
assoc_cor <- with(simpred_cor, matlm(form, dat, pred = pred))
```

```{r qq_cor}
qq_plot(assoc_cor$tab$pval)
```

## Covariance matrix among predictors

The covariance matrix is pre-defined in `matlm_sim_randpred` function
and has a simple form:

- diagonal entries are 1;
- off-diagonal entries are `rho` (equal here to `r rho`).

```{r cmat, dependson = "simpar"}
cmat <- matlm_sim_randpred(seed = seed, N = N, M = M, rho = rho, ret = "mat")
```

```{r show_cmat}
# number of predictors
M

# dimenstions of matrix
dim(cmat)

# a sub-matrix
cmat[1:5, 1:5]
```

## Covariance matrix among test statistics

It can be shown the covariance matrix among test statistics (t-test) $s_i$
is the _correlation_ matrix among predictors $x_i$:

$cov(s_i, s_j) = cor(x_i, x_j)$

This basic relationship is true for the simplest linear regression model:

$y = \mu + \beta x_i + e$

$e \sim \mathcal{N}(0, \sigma_e^2)$

In other cases, e.g. related observations, some modifications are required.

## Dummy correction of qq-plot

```{r ch_inv}
C <- matrix(rho, M, M)
diag(C) <- 1

ch <- chol(C)
ch_inv <- solve(ch)
```

```{r correct_pvals}
s_assoc <- assoc_cor$tab$zscore
s_corrected <- as.numeric(s_assoc %*% ch_inv)

pvals_corrected <- pchisq(s_corrected^2, 1, lower.tail = FALSE)
```

```{r qq_corrected}
qq_plot(pvals_corrected)
```


 
